Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-12 vesicle_code/docs/adhesion/adhesion.tex
edited conclusions

Next we conduct a numerical experiment where a vesicle doublet is placed
at the center of a fluid trap, which can be actively controlled in
microfluidic channel so that fluid trap center is effectively the
stagnation point of an extensional flow characterized by a extensional
flow rate.  At low flow rate, we find the doublet to rotate nearly
ninety degrees to align with the flow such that the long axis of the
doublet is along the divergent axis and the convergent stream is pushing
the two vesicles together. As the flow rate increases, the vesicle
doublet rotates less, and when the flow rate reaches above the critical
value, the diverging flow breaks the doublet structure by pulling the
vesicles apart.  These results indicate that it is possib le to use the
fluid trap to separate a vesicle doublet under adhesion, and thus
provide a means to probe the adhesion strength between membranes.  For a
pair of $\mu m$-sized vesicles with a bending rigidity of $\sim
10^{-19}$ $J$, an extensional flow rate of $0.5$ $s^{-1}$ is expected to
separate a vesicle doublet with reduced area of $\Delta A = 0.8$ and a
Hamaker constant $\mathcal{H}=0.7$, which corresponds to $\sim 1$ $\mu
J/m^2$.  These conditions are quite realizable in microfluidic
experiments, and we hope that our simulations will inspire microfluidic
experiments.

We also examine how adhesive interaction may lead to formation of a
vesicle doublet dynamically in flowing conditions.  We simulate two
vesicles approaching each other in a planar shear flow, and examine how
their adhesive interactions lead to doublet formation.  Once a doublet
forms, the two vesicle membranes rotate around each other as they deform
dynamically. The usual tank-treading motion of a vesicle under shear
flow is not observed in each of the two vesicles. We compute the
effective shear viscosity of a dilute suspension of vesicle doublets,
and found it to be more than twice the effective shear viscosity of a
dilute suspension of single vesicles.  Furthermore, we find that the
membrane adhesion contribution to the shear viscosity increases with
decreasing reduced area while the bending/tension contribution increases
with the reduced area.

In our formulation we did not include any electrostatic interactions
between membranes under adhesion. When the electrostatic interaction is
important, an electro-osmotic pressure in the thin film is found to be
responsible for the observed membrane
undulation~\cite{SteinkuhlerAgudo-Canalejo2016_BJ}.  In addition,
simulations presented in this work are for identical vesicles (same
reduced area, length, and bending rigidity modulus) in the doublet.
Flormann {\it et al.} demonstrated that asymmetric vesicle reduced area
may lead to various equilibrium doublet shape such as male-female,
asymmetric S-shape and parachute shape.  It is also possible that
viscosity contrast may also lead to slightly different equilibrium
doublet shape.  Future work includes three dimensional simulations,
dispersive vesicle properties (such as reduced area and bending
modulus), viscosity contrast, and effects of confinement on adhesive
interactions.

Flormann {\it et al.} studied the clustering and packing of vesicles in
free suspension~\cite{FlormannAouane2017_SciReports}.  In engineering
applications of vesicle emulsions often smaller vesicles are enclosed in
a big vesicle, as shown in Figure~\ref{fig:relaxationManyVes}, where
the left panel shows an initial condition of five vesicles suspended
inside of a larger vesicle.  The second panel from left shows the
membrane configuration and flow streamlines at an intermediate time.
There is no imposed flow, so the dynamics are governed entirely by the
adhesion, bending, and tension forces.  In the absence of confinement,
we would expect the vesicles to align themselves as a rouleaux with the
membranes forming a parachute, male-female, or S-shape
configurations~\cite{FlormannAouane2017_SciReports}.  However, because
of the bounding vesicle and the dense packing, the interior vesicles
align themselves differently.  In addition, several parts of the
bounding membrane closely follow the interior membranes in order to
minimize the adhesion energy.   As a result of force balance the
equilibrium shape of the bounding vesicle is far from a symmetric shape
while the interior smaller vesicles are of similar shape to each other.
We also plot the tension (thrid) and adhesion (fourth) of the final
equilibrium configuration.  We remark that the adhesion force is all
attractive on all the vesicle membranes while large positive membrane
tension is found only on the bounding vesicle.  In particular, the
interior vesicles form a flat contact region with a large negative
tension.  We are now conducting simulations of such vesicle
configuration under a planar shear flow.

\begin{figure}[htp]
\centering
\includegraphics[height=0.24\textwidth,trim={10cm 0cm 10cm 0cm},clip]{figs/relaxationManyVesInitialCondition.png}
\includegraphics[height=0.24\textwidth,trim={2cm 0cm 2cm 0cm},clip]{figs/relaxationManyVesStreamlines.png}
\includegraphics[height=0.24\textwidth,trim={2cm 0cm 1cm 0cm},clip]{figs/relaxationManyVesTension.png}
\includegraphics[height=0.24\textwidth,trim={2cm 0cm 2cm 0cm},clip]{figs/relaxationManyVesAdhesion.png}
\caption{\label{fig:relaxationManyVes} Five vesicles in a quiescent flow
suspended in a bounding vesicle.   All vesicles are interacting through
hydrodynamic flow and an adhesion potential with Hamaker constant
$\mathcal{H} = 0.7$ and separation distance $\delta = 0.4$.  From left
to right, the plots are the initial condition, streamlines and an
intermediate time, tension at the steady-state configuration, and the
adhesion force projected onto the outward unit normal at the
steady-state configuration.  Along the bounding vesicle, the tension is
largest and the adhesion is smallest in the regions that are
well-separated from the interior vesicles.  Similar to other
simulations, the interior vesicles are relatively flat and have a large
negative tension in the contact regions.}
\end{figure}

Recently Liu {\em et al.}~\cite{LiuChuNewbyRead2018_bioRxiv} used
immersed boundary simulations to show that, at a separation distance of
tens of nanometers, the thin film between the two membranes facilitate
the coupling between membranes via strong hydrodynamic interactions. In
particular they demonstrate numerically that the fluctuation in one
membrane is highly correlated to the other membrane without any physical
contact. It is not clear how thermal fluctuations may affect the
hydrodynamics of vesicles under adhesion. For example, does the
fluctuating hydrodynamics in the thin film between two vesicles enhance
adhesion to keep vesicles bound under linear flows?

>>>>

Next we conduct a numerical experiment where a vesicle doublet is placed
at the center of a fluid trap, which can be actively controlled in
microfluidic channel so that fluid trap center is effectively the
stagnation point of an extensional flow characterized by a extensional
flow rate.  At low flow rate, we find the doublet to rotate nearly
ninety degrees to align with the flow such that the long axis of the
doublet is parallel to the divergent axis and the convergent stream is
pushing the two vesicles together. As the flow rate increases, the
vesicle doublet rotates less, and when the flow rate exceeds the
critical value, the diverging flow breaks the doublet structure by
pulling the vesicles apart.  These results indicate that it is possible
to use the fluid trap to separate a vesicle doublet under adhesion, and
thus provide a means to probe the adhesion strength between membranes.
For a pair of $\mu m$-sized vesicles with a bending rigidity of $\sim
10^{-19}$ $J$, an extensional flow rate of $0.5$ $s^{-1}$ is expected to
separate a vesicle doublet with reduced area of $\Delta A = 0.8$ and a
Hamaker constant $\mathcal{H}=0.7$, which corresponds to $\sim 1$ $\mu
J/m^2$.  These conditions are quite realizable in microfluidic
experiments, and we hope that our simulations will inspire microfluidic
experiments.

We also examine how adhesive interaction may dynamically lead to the
formation of a vesicle doublet.  We simulate two vesicles approaching
each other in a planar shear flow, and examine how their adhesive
interactions lead to doublet formation.  Once a doublet forms, the two
vesicle membranes rotate around each other as they deform dynamically.
The usual tank-treading motion of a vesicle under shear flow is not
observed in each of the two vesicles. We compute the effective shear
viscosity of a dilute suspension of vesicle doublets, and found it to be
more than twice the effective shear viscosity of a dilute suspension of
single vesicles.  Furthermore, we find that the membrane adhesion
contribution to the shear viscosity increases with decreasing reduced
area while the bending/tension contribution increases with the reduced
area.

In our formulation we did not include any electrostatic interactions
between membranes under adhesion. When the electrostatic interaction is
important, an electro-osmotic pressure in the thin film is found to be
responsible for the observed membrane
undulation~\cite{SteinkuhlerAgudo-Canalejo2016_BJ}.  In addition,
simulations presented in this work are for identical vesicles (same
reduced area, length, and bending rigidity modulus) in the doublet.
Flormann {\it et al.} demonstrated that asymmetric vesicle reduced area
may lead to various equilibrium doublet shape such as male-female,
asymmetric S-shape, and parachute shape.  It is also possible that
viscosity contrast may also lead to different equilibrium doublet
shapes.  Future work includes three dimensional simulations, dispersive
vesicle properties (such as reduced area and bending modulus), viscosity
contrast, and effects of confinement on adhesive interactions.

Flormann {\it et al.} studied the clustering and packing of vesicles in
free suspension~\cite{FlormannAouane2017_SciReports}.  In engineering
applications of vesicle emulsions, smaller vesicles are often enclosed in
a big vesicle, as shown in Figure~\ref{fig:relaxationManyVes}, where
the left panel shows an initial condition of five vesicles suspended
inside of a larger vesicle.  The second panel from the left shows the
membrane configuration and flow streamlines at an intermediate time.
There is no imposed flow, so the dynamics are governed entirely by the
vesicle forces, including adhesion.  In the absence of confinement,
we would expect the vesicles to align themselves as a rouleaux with the
membranes forming a parachute, male-female, or S-shape
configuration~\cite{FlormannAouane2017_SciReports}.  However, because of
the bounding vesicle and the dense packing, the interior vesicles align
themselves differently.  In addition, several parts of the bounding
membrane closely follow the interior membranes in order to minimize the
adhesion energy.   As a result of force balance, the equilibrium shape
of the bounding vesicle is far from a symmetric shape while the interior
smaller vesicles are of similar shape to each other.  We also plot the
tension (third) and adhesion (fourth) of the final equilibrium
configuration.  We remark that the adhesion force is attractive on all
the vesicle membranes while large positive membrane tension is found
only on the bounding vesicle.  In particular, the interior vesicles form
a flat contact region with a large negative tension.  We are now
conducting simulations of such vesicle configuration under a planar
shear flow.

\begin{figure}[htp]
\centering
\includegraphics[height=0.24\textwidth,trim={10cm 0cm 10cm 0cm},clip]{figs/relaxationManyVesInitialCondition.png}
\includegraphics[height=0.24\textwidth,trim={2cm 0cm 2cm 0cm},clip]{figs/relaxationManyVesStreamlines.png}
\includegraphics[height=0.24\textwidth,trim={2cm 0cm 1cm 0cm},clip]{figs/relaxationManyVesTension.png}
\includegraphics[height=0.24\textwidth,trim={2cm 0cm 2cm 0cm},clip]{figs/relaxationManyVesAdhesion.png}
\caption{\label{fig:relaxationManyVes} Five vesicles in a quiescent flow
suspended in a bounding vesicle.   All vesicles are interacting through
hydrodynamic flow and an adhesion potential with Hamaker constant
$\mathcal{H} = 0.7$ and separation distance $\delta = 0.4$.  From left
to right, the plots are the initial condition, streamlines at an
intermediate time, tension at the steady-state configuration, and the
adhesion force projected onto the outward unit normal at the
steady-state configuration.  Along the bounding vesicle, the tension is
largest and the adhesion is smallest in the regions that are
well-separated from the interior vesicles.  Similar to other
simulations, the interior vesicles are relatively flat and have a large
negative tension in the contact regions.}
\end{figure}

Recently Liu {\em et al.}~\cite{LiuChuNewbyRead2018_bioRxiv} used
immersed boundary simulations to show that, at a separation distance of
tens of nanometers, the thin film between the two membranes facilitate
the coupling between membranes via strong hydrodynamic interactions. In
particular, they demonstrate numerically that the fluctuation in one
membrane is highly correlated to the other membrane without any physical
contact. It is not clear how thermal fluctuations may affect the
hydrodynamics of vesicles under adhesion. For example, does the
fluctuating hydrodynamics in the thin film between two vesicles enhance
adhesion to keep vesicles bound under linear flows?
